### 2019-08-15 response.txt
rename revision.txt

Response to the first reviewer:

- Extending to n-Type:
  As we note in section 2, generalizing external types in signatures to Type-i
  is possible and not difficult. We omit this generalization from the paper
  mainly in order to sync with the existing Agda formalization.

- Page 1: initiality vs finite generation
  We revised the wording.

- Lambdas in signatures:

  The only function space where it makes sense to have lambdas is the
  (7) infinitary function space, since we can only mention these in
  path types. But such lambdas can be always eliminated by introducing
  extra function parameters which are constrained by pointwise
  equalities. In the case of the mentioned example,

    z : H
    n : (A → H) → H
    q : n (λ x. z) = z

  We may have the following equivalent λ-free type for q:

    q : (f : A → H) → ((x : A) → f x = z) → n f = z

  In general, if we have
    Γ;Δ ⊢ λx. t : El ((x : a) → b)
  Then instead we may have
    Γ;Δ, f : El ((x : a) → b), p : (x : a) → f x = t ⊢ f : El ((x : a) → b)

  We include this explanation in the revision.

- Page 7, 8 item (5), about type equalities and UA

  We include this in the revision.

- Page 10, about initial NatAlg
  We use Nat* instead of initial Nat-algebra in the revised version.

- Page 11, about coherence issue, references

  In the revision we add a reference to the following discussion:

    https://homotopytypetheory.org/2014/03/03/hott-should-eat-itself/

  As discussed there, our coherence problem is related to the better
  known problem of constructing semisimplicial types; solving the
  former implies a solution of the latter.

- Page 12, about initial CwF

  We removed the problematic "initial" occurrence. To say a bit more
  here, if we set-truncate CwFs then there is a strict category of
  CwFs, and if we don't truncate then we conjecture that there is a
  wild category of wild CwF-s, but in any case we are referring to a
  1-category for the purpose of induction and recursion. We think that
  the subtleties involving these categories are not very relevant to
  the discussion in that section, so we avoid it in the revision.

- Page 13, "would merely decidable equality also be sufficient?"

  For the purpose of elaborating raw syntax into intrinsic syntax
  (which is in our view an important motivation for decidability), it
  would not be sufficient, because we need to eliminate from decided
  equalities to proof-relevant intrinsic terms.

--------------------------------------------------------------------------------

Response to the second reviewer:

- On the suggestion for identity homomorphisms:

  This suggestion does solve the U[] case. However, it still does not
  work; some cases (like U[]) become provable, where we can make use
  of strict equality hypotheses, but other cases fail, where we need
  to provide strict equalities. For example, assuming the suggested
  strict tᴵᴰ equations, we have

    (t =ₐ u)ᴵᴰ γ : (t = u)ᴹ γ γ (Γᴵᴰ γ) ≡ Uᴵᴰ γ ((t = u)ᴬ γ)

  The type can be reduced along definitions, and using tᴵᴰ and uᴵᴰ, to:

    (λ e. refl ◾ ap (λ x. x) e ◾ refl) ≡ (λ e. e)

  Here, e : tᴬ γ = uᴬ γ, and it is necessarily a non-strict equality. Hence,
  this case is not provable with strict tᴵᴰ equalities but provable with weak
  ones. We include this explanation in the revision.

- Addressing comments:
  - page 7, Figure 2, about external parameters:

    We do not construct HIIT definitions, only signatures. HIIT signatures can
    only depend on an HIIT specified by a previous signature if we assume the
    existence of that HIIT in the external type theory (i.e. as part of the
    context Γ). Also, the codomain types of external (6) and infinitary (7)
    functions can depend on the additional (x : A) external binder. For example,
    if we have Γ = (Nat: Type₀, zero: Nat, suc: Nat → Nat), and we write the
    signature for length-indexed A-vectors, then we have "cons : (n : Nat) → A →
    Vec n → Vec (suc n)", where the "suc n" expression refers back to the
    external "n : Nat" binder. The only way to extend the external Γ is by using
    (6) and (7) functions, otherwise, we can only refer to Γ entries which are
    assumed on the "top level", such as the A : Type₀ in the example for lists
    on page 8 in Section 2.1.

  - page 12: for the intrinsic syntax, do you also assume an induction
    principle for it, e.g. that it is initial?

    Yes. Of course, the induction principle for the intrinsic syntax would be
    something that is formalized in the current paper, so if we want to avoid
    circularity, then we need to say that the induction principle is given in an
    ad-hoc fashion. This is probably acceptable, since such ad-hoc HIT induction
    principles have been common in previous HoTT literature.

  - page 12, line -9: you say that you "need to" set-truncate; is this
    rather a strong belief that you have to?

    In the next paragraphs we justify the need for truncation. We
    think that undecidability of conversion and impossibility of
    extrinsic term models are solid justifications.

  - page 21, line -16: is there an expected equation if all computation rules
    were definitional?

    We think that if we follow "book" HoTT practice, p^s would be a
    propositional computation rule, because it involves "coe" which is defined
    externally, similarly to how path constructor computation rules involving
    "ap" are also propositional.